---
title: "Homework 9"
author: "Katie Bardsley"
date: "4/7/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#### Set up functions and global variables

```{r}
# write functions

###################################
# FUNCTION: make_dataframe
# description: take parameters and turn them into data frame 
#               of parameters to use for future tasks
# inputs: n, mean, and variance for each treatment
# outputs: dataframe
# ---------------------------------
make_dataframe <- function(n_cold, mean_cold, var_cold,
                           n_warm, mean_warm, var_warm){
  dataframe <- data.frame(n_cold, mean_cold, var_cold,
                          n_warm, mean_warm, var_warm)
  return(dataframe)
} # end of make_dataframe

###################################

###################################
# FUNCTION: RandomBudbreakData
# description: generate random budbreak data and organize in data frame
# inputs: data frame with n, mean, and variance for each treatement
# outputs: dataframe
# ---------------------------------
RandomBudbreakData <- function(data){
  data_cold <- rnorm(n=data$n_cold, mean=data$mean_cold, sd=sqrt(data$var_cold))
  data_warm <- rnorm(n=data$n_warm, mean=data$mean_warm, sd=sqrt(data$var_warm))

# combine into one big data frame
  cold <- data.frame(1:data$n_cold, data_cold, 
                     rep("cold", data$n_cold))
  colnames(cold) <- c("ID", "BudBreakDay","Climate")
  warm <- data.frame((data$n_cold+1):(data$n_cold+data$n_warm), 
                     data_warm, rep("warm", data$n_cold))
  colnames(warm) <- c("ID", "BudBreakDay","Climate")
  budbreak_data <- rbind(cold, warm)
  return(budbreak_data)
} # end of RandomBudbreakData

###################################

###################################
# FUNCTION: plot_data
# description: make boxplots for budbreak data set
# inputs: dataframe with n, mean, and variance for each treatment
# outputs: boxplot
# ---------------------------------
plot_data <- function(budbreak_data){
  ggplot2::ggplot(data=budbreak_data, mapping=
                    aes(y=BudBreakDay, fill=Climate)) +
    geom_boxplot() + 
    labs(title="Julian days before budbreak for poplar samples from different climates", 
       y="Julian days before budbreak") + 
  theme_minimal()

} # end of plot_data

###################################

###################################
# FUNCTION: anova_budbreak
# description: create anova model and return summary
# inputs: budbreak data
# outputs: summary of anova model
# ---------------------------------
anova_budbreak <- function(data){
  mod <- aov(BudBreakDay ~ Climate, data=data)
  return(summary(mod))
} # end of anova_budbreak

###################################

```


```{r}
# set global variables

# cold source climate
n_cold <- 500
mean_cold <- 125
var_cold <- 300
  
# warm source climate
n_warm <- 500
mean_warm <- 140
var_warm <- 300
```


#### 1. Creating a random bud break data set and run ANOVA analysis

Here I will be focusing on bud flush phenology in poplar, thinking about patterns similar to those being studied in the Keller Lab currently. In a common garden, you might expect samples taken from colder climates to break bud earlier than those from warmer climates as they have a shorter growing season and need to take advantage of all the warm days they can get. 


```{r}
budbreak <- RandomBudbreakData(make_dataframe(n_cold, mean_cold, var_cold, 
                                              n_warm, mean_warm, var_warm)) 
              # make random data set
anova_budbreak(budbreak) # make anova model and produce summary
plot_data(budbreak) # plot DENSITY plots of data
```

#### 2. Creating a new function

```{r}

###################################
# FUNCTION: plot_data_density
# description: make density plots for budbreak data 
# inputs: dataframe with n, mean, and variance for each treatment
# outputs: density plots
# ---------------------------------
plot_data_density <- function(budbreak_data){
  ggplot2::ggplot(data=budbreak_data, mapping=
                    aes(x=BudBreakDay, fill=Climate)) +
    geom_density(alpha=0.8) + 
    labs(title="Julian days before budbreak for poplar samples from different climates", 
       x="Julian days before budbreak") + 
  theme_minimal()

} # end of plot_data_density

###################################

```

```{r}

budbreak <- RandomBudbreakData(make_dataframe(n_cold, mean_cold, var_cold, 
                                              n_warm, mean_warm, var_warm)) 
              # make random data set
anova_budbreak(budbreak) # make anova model and produce summary
plot_data_density(budbreak) # plot boxplots of data

```


#### 3. Changing the sample sizes within a for loop and recording p-values

Play with different sample sizes

```{r}

# try playing with different sample sizes
sample_sizes <- c(10, 20, 30, 40, 50, 100, 200, 400, 500, 1000)
mean <- rep(999, length(sample_sizes))
count <- 1

for (n in sample_sizes){
  n_warm <- n/2 ; n_cold <- n/2
  pval <- rep(999, 100)
  for (i in 1:100){
    budbreak <- RandomBudbreakData(make_dataframe(n_cold, mean_cold, var_cold, 
                                                  n_warm, mean_warm, var_warm)) 
    # make random data set
    summary <- anova_budbreak(budbreak) # make anova model and produce summary
    pval[i] <- summary[[1]]$`Pr(>F)`[1]  # save p-value
  }
  mean[count] <- signif(mean(pval), digits=3)
  count <- count + 1
  
}
means <- data.frame(sample_sizes/2, mean)
colnames(means) <- c("Number of samples in each treatment", 
                     "Mean p-value across 100 iterations")
print(means)

```



